---
title: "Data Analysis Exercise"
output:
  html_document:
    toc: FALSE
editor: 
  markdown: 
    wrap: 72
---

# Data Information

This data set contains the deaths involving COVID-19, pneumonia, and
influenza reported to NCHS by sex, age group, and jurisdiction

## Data Source

This data was obtained from the [Centers for Disease Control and
Prevention (CDC)'s National Center for Health Statistics
(NCHS)](https://data.cdc.gov/NCHS/Provisional-COVID-19-Deaths-by-Sex-and-Age/9bhg-hcku).

## Data Contents

This dataset contains 116,000 rows and 16 columns.

# Data Loading and Processing

## Load Libraries

```{r echo = FALSE}
# Load Libraries
library(tidyverse)
library(gtsummary)
library(lubridate)
```

## Load Data

```{r}
Raw_Data <- read_csv("data-analysis-exercise/data/Provisional_COVID-19_Deaths_by_Sex_and_Age.csv")
```

## Data Contents

```{r}
str(Raw_Data)
```

# Cleaning Data of interest

Remove special characters from column names

```{r}

# Save copy of the raw data to clean
data = Raw_Data

#remove special characters
names(data) = gsub(" ", "_",
           gsub("-", "_",
           gsub(",", "", names(data))))
```

**Change data types where necessary.**\
character -\> date\
numeric -\> character

```{r}
data = data %>% 
  mutate_at(
    vars('Data_As_Of', 'Start_Date','End_Date'), 
    as_date,
    format = "%m-%d-%y"
    )%>% mutate(Year = as.character(Year))
```

## Determine what data to keep

```{r}
dfm = data %>% pivot_longer(cols=c('COVID_19_Deaths','Pneumonia_Deaths','Influenza_Deaths'),
                    names_to="variable",
                    values_to="value")
year_plot = ggplot(dfm,aes(x = Year,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge")+theme(legend.position = "bottom")

age_plot = ggplot(dfm,aes(x = Age_Group,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge")+theme(legend.position="none",axis.text.x = element_text(size = 10, angle=60, hjust=1))

grid.arrange(year_plot, age_plot, ncol=2)

```

According to the footnotes column, there are multiple data points that
had been suppressed due to NCHS confidentiality standards. Separated by
state and/or age have many points that have been suppressed. The best
option is to subset the data by year, and keep only rows containing all
ages, and states.

**Subset to keep**\
collected in 2020 and 2021

```{r}
data = data %>% filter(State == "United States",Age_Group=="All Ages",Year == 2020 | Year == 2021, Sex == "All Sexes")%>%
  select(Start_Date, End_Date, COVID_19_Deaths,Pneumonia_Deaths,Influenza_Deaths, Group, Sex, Year)
```

```{r}
data
```

# Save Data and Summary

```{r}
#summary(data)
summary_df = data.frame(do.call(cbind, lapply(data, summary)))
print(summary_df)
```


```{r}
str_data = str(data)
summary_data = summary(data)

summary_data


```


# Save RDS File

```{r}
saveRDS(data, file = "data-analysis-exercise/clean/clean_data.rds")
saveRDS(summary_data, file = "data-analysis-exercise/output/data_summary.rds")
```
