---
title: "Flu Anlaysis - Wrangling"
output:
  html_document:
    toc: FALSE
editor: 
  markdown: 
    wrap: 72
---

# Data Loading and Processing

## Load Libraries

```{r echo = FALSE}
# Load Libraries
library(tidyverse)
library(dplyr)
library(here) 
library(skimr)
library(tidyr)
library(tidymodels)
```

## Load the data

```{r}
raw_data = readRDS(here("fluanalysis", "data", "SympAct_Any_Pos.Rda")) #load RDS file
```

## Data Contents

```{r}
str(raw_data) #ensure the data is complete
```

## Process data as follows:

**1. Remove all variables that have *Score* or *Total* or *FluA* or
*FluB* or *Dxname* or *Activity* or *Unique.Visit***

**2. Remove all observations with NA**

Don't do this manually one by one, figure out how to use R commands that
let you remove things in an efficient manner.

```{r}
clean_data = raw_data %>% #create new variable to ensure raw_data is not manipulated
  select(-contains(c("Score","Total","FluA","FluB","Dxname","Activity")))%>% #select columns to remove
  select(-c('Unique.Visit')) %>% #remove collumn
  na.omit() #remove all observations with NA
```

## Additional Processing (added for machine learning modeling Module 11)

**remove yes/no versions for symptoms with multiple levels**
```{r}
clean_data_update =
  clean_data %>%
  select(!c(WeaknessYN, CoughYN, MyalgiaYN, CoughYN2))
```


**categorical/ordinal predictors**
there are categorical and ordinal predictors. Code categorical variables as unordered factors and others as ordered factors.
```{r}
# categorical productors 
unorderedRecipe = recipe(~ SwollenLymphNodes + ChestCongestion + ChillsSweats + NasalCongestion + Sneeze + Fatigue + SubjectiveFever + Headache + RunnyNose + AbPain + ChestPain + Diarrhea + EyePn + Insomnia + ItchyEye + Nausea + EarPn + Pharyngitis + Breathless + ToothPn + Vomit + Wheeze, data = clean_data_update)

unorderedTest = unorderedRecipe %>%
  step_dummy(all_predictors()) %>%
  prep(training = clean_data_update)

categoricalTestData = bake(unorderedTest, new_data = NULL)
```

```{r}
#ordinal predictor

#specify the ordinal levels as found in the data
ordinalLevels = c("None", "Mild", "Moderate", "Severe")
clean_data_update = clean_data_update %>%
  mutate(Weakness = ordered(Weakness),
         CoughIntensity = ordered(CoughIntensity),
         Myalgia = ordered(Myalgia))

#create recipe (like categorical above)
ordinalRecipe = recipe(~ Weakness + CoughIntensity + Myalgia, data = clean_data_update)
ordinalTest = ordinalRecipe %>%
  step_ordinalscore(all_predictors()) %>%
  prep(training = clean_data_update)
ordinalTestData <- bake(ordinalTest, new_data = NULL)
```


**identify and remove binary predictors that have <50 entries in one category**
```{r}
summary(clean_data_update)
# vision and hearing have les than 50

clean_data_update = 
  clean_data_update %>%
  select(!c(Vision, Hearing))

```


## Save RDS File

```{r}
#save cleaned data
cleandata_location = here("fluanalysis", "data", "cleandata.rds")
saveRDS(clean_data, file = cleandata_location)
```

```{r}
# save cleaned data after Module 11
cleandata_location2 = here("fluanalysis", "data", "cleandata2.rds")
saveRDS(clean_data_update, file = cleandata_location2)
```

